---
title: "exercise2"
author: "Valerie Arnaldi"
date: "6 5 2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Load the necessary libraries ###
```{r}
library(tidyverse) 
library(readr)
library(SimilarityMeasures)
library(lubridate)  
library(cowplot)      
```

## Import data
```{r}
wildschwein_BE <- read_delim("wildschwein_BE_2056.csv",",")
```
# filter wildboar Sabi
```{r}
sabi <- filter(wildschwein_BE,TierName == "Sabi")
```
# Step b) Measure the distance from every point to every other point within the temporal window of 60 min
```{r}
sabi <- sabi %>%
  mutate(
    nMinus2 = sqrt((lag(E,2)-E)^2+(lag(N,2)-N)^2),   # distance to pos -30 minutes
    nMinus1 = sqrt((lag(E,1)-E)^2+(lag(N,1)-N)^2),   # distance to pos -15 minutes
    nPlus1  = sqrt((E-lead(E,1))^2+(N-lead(N,1))^2), # distance to pos +15 mintues
    nPlus2  = sqrt((E-lead(E,2))^2+(N-lead(N,2))^2)  # distance to pos +30 minutes
  )
```
# Calculate the mean distance of nMinus2, nMinus1, nPlus1, nPlus2
# Since we want the mean value per Row, we have to explicitly specify this before mutate() with the function rowwise()
# To remove this rowwise-grouping, we end the operation with ungroup()
# Note that for the first two positions, we cannot calculate a stepMean since there is no Position n-2 for these positions.
# This is also true for the last to positions (lacking a position n+2).
```{r}
sabi <- sabi %>%
  rowwise() %>%
  mutate(
    stepMean = mean(c(nMinus2, nMinus1,nPlus1,nPlus2))
  ) %>%
  ungroup() 

sabi
```
# Step c): Remove "static points
# We can now determine if an animal is moving or not by specifying a threshold distance on stepMean
# In our example, we use the mean value as a threshold:
# Positions with distances below this value are considered static
```{r}
sabi <- sabi %>% 
  ungroup() %>%
  mutate(static = stepMean < mean(stepMean, na.rm = TRUE)) # "stepMean" muss mit Zahl als Treshold ersetzt werden

sabi_filter <- sabi %>%
  filter(!static)

sabi_filter%>%
  ggplot(aes(E, N))  +
  geom_path() +
  geom_point() +
  coord_fixed() +
  theme(legend.position = "bottom")
```

### Task 1
# Segmentation
# Import Data
```{r}
Caro <- read_delim("caro60.csv",",")
```
# The sampling interval for this dataset is 1 minute
# Use a temporal window v of 6 minutes
```{r}
Caro <- Caro %>%
  mutate(
    nMinus3 = sqrt((lag(E,3)-E)^2+(lag(N,3)-N)^2),   # distance to pos -3 minutes
    nMinus2 =
sqrt((lag(E,2)-E)^2+(lag(N,2)-N)^2),
 # distance to pos -2 minutes
    nMinus1 = sqrt((lag(E,1)-E)^2+(lag(N,1)-N)^2),   # distance to pos -1 minute
    nPlus1  = sqrt((E-lead(E,1))^2+(N-lead(N,1))^2), # distance to pos +1 mintue
    nPlus2  = sqrt((E-lead(E,2))^2+(N-lead(N,2))^2),  # distance to pos +2 minutes
    nPlus3  =
sqrt((E-lead(E,3))^2+(N-lead(N,3))^2)
  )
 # distance to pos +3 minutes
```

### Task 2
# Specify and apply treshold d
```{r}
Caro <- Caro %>%
  rowwise() %>%
  mutate(
    stepMean = mean(c(nMinus3, nMinus2, nMinus1,nPlus1,nPlus2,nPlus3))
  ) %>%
  ungroup() 

Caro
```
# Summary statistics
```{r}
summary(Caro$stepMean)
boxplot(Caro$stepMean)
hist(Caro$stepMean)
# die meisten Werte liegen zwischen 5-10
```
# define reasonable threshold value to differentiate between stops and moves
```{r}
Caro <- Caro %>% 
  ungroup() %>%
  mutate(static = stepMean < mean(5, na.rm = TRUE))

Caro_filter <- Caro %>%
  filter(!static)

sabi_filter%>%
  ggplot(aes(E, N))  +
  geom_path() +
  geom_point() +
  coord_fixed() +
  theme(legend.position = "bottom")
```

### Task 3
# Visualize segmented trajectories

```{r}
Caro%>%
  ggplot(aes(E, N, color = static))  +
  geom_path() +
  geom_point() +
  coord_fixed() +
  theme(legend.position = "bottom")
```

### Task 4
# Segment-based analysis
# In order to complete the last steps (c and d), we need a unique ID for each segment that we can use as a grouping variable.
```{r}
rle_id <- function(vec){
  x <- rle(vec)$lengths
  as.factor(rep(seq_along(x), times=x))
  }
```
# You can use the newly created function rle_id to assign unique IDs to subtrajectories
```{r}
Caro <- Caro %>%
  mutate(segment_id = rle_id(static))
```
# Visualize the moving segments by colourizing them by segment_ID
```{r}

Caro_filter <- filter(Caro,static == FALSE)

Caro_filter%>%
  ggplot(aes(E, N, color = segment_id))  +
  geom_path() +
  geom_point() +
  coord_fixed() +
  theme(legend.position = "bottom")+
 labs(title = "All segments (uncleaned)")

Caro_Long <- Caro_filter %>%
  group_by(segment_id) %>%
  summarize(
    duration = max(DatetimeUTC)-min(DatetimeUTC)
  )

Caro_Join <- full_join(Caro_filter, Caro_Long, by = "segment_id")

Caro_Join_Filter <- filter(Caro_Join,duration > 300)

Caro_Join_Filter%>%
  ggplot(aes(E, N, color = segment_id))  +
  geom_path() +
  geom_point() +
  coord_fixed() +
  theme(legend.position = "bottom")+
 labs(title = "Long segments (removed segments <5)")
```

### Task 5
# Similarity measures on human movement
# Import Data
```{r}
pedestrian <- read_delim("pedestrian.csv",",")
```
# explore the trajectories first and get an idea on how the pedestrians moved
```{r}
pedestrian%>%
  ggplot(aes(E, N, color = TrajID))  +
  geom_path() +
  geom_point() +
  coord_fixed() +
  theme(legend.position = "top")+
 labs(title = "Visual comparison of the 6 trajectories", subtitle = "Each subplot highlights a trajectory") +
theme(legend.position = "none")+
  facet_wrap(~ TrajID, nrow = 2)
```

### Task 6
# Calculate similarity
# install package SimilarityMeasures
```{r}
# install.packages("SimilarityMeasures")
# library("SimilarityMeasures")
```
# get familiar with the package
```{r}
help(package = "SimilarityMeasures")
```
# select and safe single trajectories
```{r}
Traj1 <- pedestrian %>%
  filter(TrajID == 1)
Traj2 <- pedestrian %>%
  filter(TrajID == 2)
Traj3 <- pedestrian %>%
  filter(TrajID == 3)
Traj4 <- pedestrian %>%
  filter(TrajID == 4)
Traj5 <- pedestrian %>%
  filter(TrajID == 5)
Traj6 <- pedestrian %>%
  filter(TrajID == 6)
```
# change from dataframe to matrix
```{r}
Traj1_mtx <- data.matrix(Traj1)
Traj2_mtx <- data.matrix(Traj2)
Traj3_mtx <- data.matrix(Traj3)
Traj4_mtx <- data.matrix(Traj4)
Traj5_mtx <- data.matrix(Traj5)
Traj6_mtx <- data.matrix(Traj6)

```
# Calculate the dynamic time warping value between two trajectories
```{r}
TJ2 <- DTW(Traj1_mtx, Traj2_mtx, 4)
TJ3 <- DTW(Traj1_mtx, Traj3_mtx, 4)
TJ4 <- DTW(Traj1_mtx, Traj4_mtx, 4)
TJ5 <- DTW(Traj1_mtx, Traj5_mtx, 4)
TJ6 <- DTW(Traj1_mtx, Traj6_mtx, 4)
```
# Calculate Edit Distance
# EditDist (traj1, traj2, pointDistance = 20)
# point distance: floating point number representing maximum distance in each dimension allowed for points to be considered equivalent
```{r}
TJ2_editdist <- EditDist(Traj1_mtx, Traj2_mtx, pointDistance = 4)
TJ3_editdist <- EditDist(Traj1_mtx, Traj3_mtx, pointDistance = 4)
TJ4_editdist <- EditDist(Traj1_mtx, Traj4_mtx, pointDistance = 4)
TJ5_editdist <- EditDist(Traj1_mtx, Traj5_mtx, pointDistance = 4)
TJ6_editdist <- EditDist(Traj1_mtx, Traj6_mtx, pointDistance = 4)
```
# Calculate Frechet
# Frechet (traj1, traj2, testLeash = -1)
# testleash: numeric value, which if positive, checks whether the leash can be used -> results in boolean; if values negative, then it is not used and the standard calculation is performed
```{r}
TJ2_frechet <- Frechet(Traj1_mtx, Traj2_mtx, testLeash = -1)
TJ3_frechet <- Frechet(Traj1_mtx, Traj3_mtx, testLeash = -1)
TJ4_frechet <- Frechet(Traj1_mtx, Traj4_mtx, testLeash = -1)
TJ5_frechet <- Frechet(Traj1_mtx, Traj5_mtx, testLeash = -1)
TJ6_frechet <- Frechet(Traj1_mtx, Traj6_mtx, testLeash = -1)
```
# Calculate LCSS (allowing translations)
# LCSS (traj1, traj2, pointSpacing = -1, pointDistance = 20, errorMarg = 2, returnTrans = FALSE)
# pointSpacing: integer value of the maximum index difference between trajectory1 and trajectory2 allowed in the calculation. A negative value sets the point spacing to unlimited.
# pointDistance: floating point number representing the maximum distance in each dimension allowed for points to be considered equivalent.
# errorMarg: floating point error margin used to scale the accuracy and speed of the calculation.
#returnTrans: boolean value to allow the best translation found to be returned as well as the LCSS value if set to true
```{r}
TJ2_lcss <- LCSS(Traj1_mtx, Traj2_mtx, pointSpacing = 4, pointDistance = 4, errorMarg = 2, returnTrans = FALSE)
TJ3_lcss <- LCSS(Traj1_mtx, Traj3_mtx, pointSpacing = 4, pointDistance = 4, errorMarg = 2, returnTrans = FALSE)
TJ4_lcss <- LCSS(Traj1_mtx, Traj4_mtx, pointSpacing = 4, pointDistance = 4, errorMarg = 2, returnTrans = FALSE)
TJ5_lcss <- LCSS(Traj1_mtx, Traj5_mtx, pointSpacing = 4, pointDistance = 4, errorMarg = 2, returnTrans = FALSE)
TJ6_lcss <- LCSS(Traj1_mtx, Traj6_mtx, pointSpacing = 4, pointDistance = 4, errorMarg = 2, returnTrans = FALSE)
```
# Create Tibble/Dataframe from calculated values
```{r}
dtw <- c(TJ2, TJ3, TJ4, TJ5, TJ6)
editdist <- c(TJ2_editdist, TJ3_editdist, TJ4_editdist, TJ5_editdist, TJ6_editdist)
frechet <- c(TJ2_frechet, TJ3_frechet, TJ4_frechet, TJ5_frechet, TJ6_frechet)
lcss <- c(TJ2_lcss, TJ3_lcss, TJ4_lcss, TJ5_lcss, TJ6_lcss)

Traj_all <- tibble(dtw, editdist, frechet, lcss, trajectory = c(2,3,4,5,6))
```
# Umformen des Tibbles (oder Dataframes)
```{r}
Traj_long <- Traj_all %>%
  pivot_longer(names_to = "name", values_to = "value", -trajectory)
```
# Visualize and compare Traj1 to Traj2-6 using different similarity measures
# Prepare single plots for each measure method
```{r}
dtw_plot <- Traj_long %>%
  filter(name == "dtw") %>%
  ggplot()+
  theme(legend.position = "none")+
  labs(x="",title = "DTW")+
  geom_col(mapping = aes(x = trajectory, y = value, fill = trajectory))

editdist_plot <- Traj_long %>%
  filter(name == "editdist") %>%
  ggplot()+
  theme(legend.position = "none")+
  labs(x="",title = "Edit Distance")+
  geom_col(mapping = aes(x = trajectory, y = value, fill = trajectory))

frechet_plot <- Traj_long %>%
  filter(name == "frechet") %>%
  ggplot()+
  theme(legend.position = "none")+
  labs(x="",title = "Frechet")+
  geom_col(mapping = aes(x = trajectory, y = value, fill = trajectory))

lcss_plot <- Traj_long %>%
  filter(name == "lcss") %>%
  ggplot()+
  theme(legend.position = "none")+
  labs(x="",title = "LCSS")+
  geom_col(mapping = aes(x = trajectory, y = value, fill = trajectory))
```
# Zusammenfuegen aller Plots
```{r}
plot_grid(dtw_plot, editdist_plot, frechet_plot, lcss_plot)
```



